### 2017-05-23 manuscript/manuscript.Rmd

Ecological interactions often exert important selective pressures on the species
involved. For example, the phenologies of lodgepole pines and red crossbills
respond spatially to the presence of squirrels [@benk03rsc]. Likewise, palm
species undergo changes in seed morphology in response to the extinction of
birds dispersing their seeds [@gale13feb; @john17blc]. Interactions can be lost,
too, when phenologies of the species involved shift [@raff15psf]. Interactions
are, in fact, so important that the existence of a species has been inferred by
the fact that another species bore traits that matched no other known species:
@krit91dmh relates the discovery of the moth *Xanthopan morganii*, with a
proboscis famously over a foot long, which Darwin predicted would exist based
solely on the phenology of local plant *Angraecum sesquipedale*. In addition,
interactions and the emergent structures they define are distributed in similar
ways across communities at both large or small scales [@jord03ipc]. Together,
these observations suggest that much ecological structure could be the end
result of (co)evolutionary dynamics between species [@eklo12reh; @stou12ecs].
Unfortunately, although the evolutionary dynamics of pairs of interacting
species have been well described at macro-evolutionary [@van73nel] and
micro-evolutionary [@gand08hcp] timescales, most attempts to understand how they
cascade up to the levels of diversity of both species and interactions found
within empirical communities have been inconclusive [@hemb14cdl]. This suggests
that these well-described mechanisms may not confer substantial predictive power
when examined at scales of organization larger than the pairwise interaction.

Historically, the evidence for shared evolutionary history in taxonomically
diverse communities relied on the quantification of the degree of matching
between the phylogenies of two sets of interacting organisms, accounting for the
distributions of interactions across the phylogeny [@lege02sth]. This notion
builds on the century-old idea that extant species interact in a way similar to
the way their ancestors did [@fahr13eua; @nuis13cam; @guim11ecm]. Note that
testing these assumptions is related to, but markedly more restrictive than,
testing for phylogenetic conservatism of species' interactions [@reze07ncp;
@eklo12reh]. This is because of additional, higher-order constraints related to
the shape of both trees at *all* depths [@cave09mce; @mouq12eap], because
ancestral evolutionary innovations have a high phylogenetic inertia, and they
carry forward to extant taxa [@vale10cpr; @desd03qps; @dini08mgc]. In a way, the
true measure of phylogenetic signal of interactions should depend not only on
how they are conserved within the tree of the species establishing them (*e.g.*
parasites or pollinators), but also how these interactions at matched to the
tree of the species receiving them (*e.g.* hosts or plants). This is true
whether or not the species complex are coevolving: in fact, neutral interactions
can yield perfectly matching co-phylogenies [@pois15wce]. Consequently, many of
the systems that have been described as exhibiting significant phylogenetic
structure of interactions ultimately deviate from this last constraint, and this
can occur for a variety of factors that stem from how other species evolved and
established, lost, or maintained interactions throughout their joint
evolutionary history (in addition to the fact that species may have different
impacts on one another, *i.e.* the effect of any interaction on the
community-level co-phylogenetic signal is not expected to be equal). Nonetheless,
detecting matching phylogenies for interacting clades indicates that their
shared evolutionary history is long standing and is therefore suggestive that
their extant ecological structure is an outcome of ancestral constraints and/or
co-adaptation [@nuis15pri].

It is important to note further that discovering matching phylogenies does not
mean that coevolutionary dynamics---*sensu* @thom99rmc---took place at any time.
In fact, coevolution is not expected to necessarily result in matching
phylogenies nor are matching phylogenies only produced through coevolution
[@pois15wce]. It follows that community-level measures of phylogenetic signal,
while they *do* quantify how closely interactions are a product of phylogeny, do
not allow us to draw conclusions on coevolution. Nevertheless, interaction-level
measures are useful in that, when expressed as the contribution of interactions
to the overall signal, they allow us to *compare* the importance of interactions
across replicated communities. Communities from the same regional pool vary
because (i) the local species pool is at best a subset of the regional species
pool and (ii) the local interactions are at best a subset of the interactions in
the regional community [@pois15swe]. This implies that (i) the phylogenetic
signal in the regional pool will be different from the signal in the local
communities, and (ii) the phylogenetic signal across local communities will
differ. Species sampling and variability of interactions, however, do no
predict (i) how the phylogenetic signal of pairwise interactions is kept or lost
at the scale of the whole community nor (ii) whether or not this variability is
related to changes in the amount of phylogenetic signal that can be detected
locally.

In this manuscript, we analyze a large dataset of over 300 species of mammalian
hosts and their ectoparasites, sampled throughout Eurasia, for which
phylogenetic relationships are known. Using a Procrustean approach to quantify
the strength of co-phylogenetic matching of interactions between host and
parasite trees [@balb13pnp], we show that locally sampled communities rarely
show strong matching despite the fact that the overall system does at the
continental scale. We then provide evidence to support the conclusion that the
amount of phylogenetic matching within a local community is predictable based on
the importance of interactions in the regional network. We finally show that the
contribution of specific interactions to phylogenetic matching is invariant
across scales, and is unrelated to their tendency to vary across space. The lack
of co-phylogenetic structure in local communities suggests that, while
interactions are undeniably important for community assembly, they might be less
so than abiotic factors.

# Methods

## Data source and pre-treatment

We use data on observations of interactions between 121 species of rodents and
205 species of parasitic fleas in 51 locations across Europe [@kras12dfp] to
build 51 species-species interaction networks. Interactions were measured by
combing rodents for fleas, a method that gives high quality data as it has a
high power of detection. The dataset also includes phylogenies for the hosts and
the parasites. Previous analyses revealed that this dataset shows significant
co-phylogenetic matching at the continental level [@kras12psm]. Importantly, it
also provides spatial replication and variability [@cana14een] at a scale large
enough to capture macro-ecological processes. This dataset is thus uniquely
suited for our analysis as it represents a thorough spatial and taxonomic
sampling of a paradigmatic system in which interspecific interactions are
thought to be driven by macro-evolution and co-speciation events [@comb01pee;
@vern09lpf].

The original dataset gives quantitative interaction strengths (expressed as an
averaged number of parasites per species per host). In this system, quantitative
interaction strengths were previously shown to be affected to a very high degree
by local variations in abundance across sampling locations [@cana14een], and it
therefore seems unlikely that they reflect macro-ecological processes.
Therefore, to account for differential sampling effort---which cannot readily be
quantified---and across site variations in abundance---which do not pertain to
macro-evolutionary processes---we only study the networks' bipartite incidence
matrices (that is, presence and absence of infection of hosts by the parasites).

## Spatial scales and interaction spatial consistency

Noting that variation of interactions across locations---which can be caused by
local ecological mechanisms as opposed to reflecting evolutionary dynamics---can
decrease congruence, we analyze the data at three different levels which we will
refer to as continental, regional, and local. Notably, the continental level
summarizes the complete dataset whereas both the regional and local levels are
location-specific scales.

*Continental* interaction data consists of the aggregated "metanetwork" which
includes all documented interactions between species from the regional species
pool [@pois12dsi]. For every location, we further define two scales of
analysis.

First, *regional* interaction data accounts for different species composition
across sites: the species that have been observed locally interact as they would
do in the *continental* network, *i.e.* if interactions did not vary across
space. This allows testing whether sampling from the regional species pool
affects co-phylogenetic matching, regardless of the distribution of
interactions. Within each site, the regional scale is given by the subset of the
metanetwork formed by the locally present species (properly speaking, the
induced subgraph of the metanetwork induced from the nodes of the local
network). Hence the regional networks are always a perfect subset of the
continental network, and do not reflect whether species were actually observed
to interact locally or not, but whether they *can* interact at all. This
*regional* network is thus a baseline estimate derived from interactions within
the species pool and measures the effect of species sampling on co-phylogenetic
matching.

Second, *local* interaction data are the actual observations at this location:
the identity of the species present, and the way they interact. In addition to
capturing the dissimilarity of species composition across sites, this allows
measuring the effect of interaction turnover across space. The local and
regional networks always include the same species, but the local network has
only a subset (or, at most, an exact match) of the interactions in the regional
network.

We finally define the spatial consistency of every interaction as the proportion
of sites in which the two co-occurring species interact with each other, or
simply

$$ S_{ij} = \frac{L_{ij}}{C_{ij}}\,. $$ {#eq:linktunover}

The spatial consistency of an interaction $S_{ij}$ between species $i$ and $j$
is therefore the ratio between the the number of locations in which they were
observed to interact ($L_{ij}$) and the number of locations in which both were
observed to be present ($C_{ij}$). Because $L_{ij} \in [0,C_{ij}]$, this measure
takes values in $[0,1]$ where larger values reflect high spatial consistency. Note
that although they are reported as 0 (*i.e.* having no interactions), we
actually have no information about species pairs that have never co-occurred;
this is a common, but hard-to-correct-for, feature of spatially replicated
datasets in which species occurrence varies [@mora15ibi]. Therefore, the only
values of $S_{ij}$ can be properly estimated are those for species pairs that
have been observed to *co-occur* at least once.

## Quantifying co-phylogenetic matching

We quantify the strength of co-phylogenetic matching in terms of the degree of
matching between host and parasite phylogenies given knowledge of extant species
interactions. We do so using the *PACo* method [@balb13pnp], which is robust to
variations in both number of species and interactions. *PACo* provides measures
of both the network-level congruence (*i.e.*, is there phylogenetic signal in
the species interactions across the entire network?) and the interaction-level
signal (*i.e.*, what is the contribution of each interaction to the overall
signal?). Because interaction-level measures provided by *PACo* operate like
residuals, larger values of this metric reflect *low* contributions to
co-phylogenetic matching. Likewise, interactions that contribute strongly to
phylogenetic congruence have smaller *PACo* values. Importantly, and in contrast
to previous methods such as *ParaFit* [@lege02sth], *PACo* also can be used to
meaningfully quantify the contribution of every interaction to the network-level
signal even in cases where the entire network shows no significant phylogenetic
signal.

All values returned by *PACo* are tested for deviation from a random expectation
[@hutc17pip], and we generated those random expectations by applying
permutations to the species interaction networks. Specifically, we applied
permutations that maintained the number of parasites for each host and the
number of hosts for each parasite. This has the effect of measuring whether
re-distributing interactions between tree tips would give rise to the same
value. We always compared the observed value to the randomized distribution
using a two-tailed statistic; thus, a significant value indicates that the
observed value is unlikely to have been observed by chance, without
pre-specifying whether or not it is larger or smaller than expected.

In PACo, the effective sample size is the number of interactions in the network,
and our interpretation of PACo's output must account for this. This is not an
issue for permutation tests, since they evaluate the significance of the
co-phylogenetic signal by permutations of each network, the power of each test
varies but the test statistics can be compared. To ensure that values of the
interaction contribution to co-phylogenetic signal are comparable, we normalized
them network-wise by dividing them by the maximal value of the sum of squares in
PACo. While the raw values returned by PACo are not meaningfully comparable
between networks, the corrected values presented here are.

As required by *PACo*, the phylogenetic trees for hosts and parasites were
rendered ultrametric (*i.e.*, all species are at the same distance from the
root). This has the consequence of losing the temporal component of the tree
(which was not available for the parasites in the original dataset), but
standardizes phylogenetic distances in a way that satisfies *PACo*'s
requirements. Moreover, this introduces the hypothesis that the common ancestor
to the parasites was able to infect the common ancestor of the host. Note that
the same procedure was applied in the original publication based on these
phylogenetic data [@kras12psm].

```{r maps, fig.show='hide', dev='pdf', fig.height=4.1, echo=F, results='hide', warning=F, message=F, error=F}
library(rworldmap)
library(igraph)
load("data/metrics.Rdata")
load("data/paco_fig1.Rdata")
sites <- read.table("data/coordinates.csv", h=T, sep="\t")
pp <- list(x=sites$Long, y=sites$Lat)
map <- getMap(resolution="low")
par(mar=c(0,0,0,0))
plot(map, xlim=range(pp$x), ylim=range(pp$y), col="#f4f2ef", border=rgb(0.87, 0.81, 0.81))

pchs <- c(21,25)[1+(fig1dat$loc < 0.05)]
bgs <- c("white", "black")[1+(fig1dat$reg < 0.05)]

points(pp$x, pp$y, pch=pchs, bg=bgs, cex=0.9)

legend("topright", c("Significant", "Non significant"), pch=c(25,21), bty="n", pt.cex=1.2, title="Local network", col="black", pt.bg=c("white", "white"))
legend("topleft", c("Significant", "Non significant"), fill=c("black", "white"), bty="n", pt.cex=0.9, title="Regional network")
box()
```

![Spatial distribution of co-phylogenetic matching across the 51 sites. For each location, we indicate whether or not the structure of regional and local interaction networks is consistent with phylogenetic congruence. Locations where the local network shows a significant co-phylogenetic structure are marked with triangles. Locaitions where the regional network shows a significant co-phylogenetic structure are in black. See also {@tbl:contingency}.](figure/maps-1.pdf){#fig:maps width=100%}

# Results and discussion

Splitting the datasets at the continental, regional, and local levels delineates
clear quantitative predictions. At the regional scale, one can expect community
assembly to promote the co-occurrence of evolutionarily linked species pairs --
*i.e.*, a host and a parasite from lineages that interact will tend to co-occur
more often because the parasites are filtered to be present in sites where they
can find hosts. Under this situation, we expect that regional networks will have
a high degree of phylogenetic matching (because they account for the information
on potential species interactions); we do in addition expect that their
phylogenetic signal will be larger than what is found in the continental
network, since the latter represents a somewhat artifactual agglomeration of
species pairs that do not co-occur. The opposite situation (a relatively lower
phylogenetic matching) would therefore be suggestive of a weaker selection for
the co-occurrence of evolutionarily tied species pairs.

At the local scale, if interactions between species at matching phylogenetic
positions are conserved, we would expect both a similar or higher level of
phylogenetic matching between the local and the regional scale, and a positive
relationship between the frequency of interaction and its overall importance for
phylogenetic matching (interactions with a strong phylogenetic signal happen
more often). On the contrary, if local assembly proceeds largely independently
from the coevolutionary history, the relative level of phylogenetic matching in
local networks should be the same as in the regional networks (through a
sampling effect from the distribution of interaction-level contribution to
co-phylogenetic matching), but the frequency of interactions should bear no
relationship to their importance in overall matching.

## Local and regional scale networks show no co-phylogenetic matching

As host-macroparasite interactions are hypothesized to be ecologically
constrained due to their being evolutionary conserved [@comb01pee], the
congruence observed at the continental level sets the baseline for what would be
expected in local communities. Of course, if ecological mechanisms (such as
filtering) reduce co-phylogenetic matching, we should detect this signal at the
continental scale but not locally. Out of 51 sites, our *PACo* analysis
indicates that 35 show no signal of co-phylogenetic matching at all, 11 show
significant co-phylogenetic matching when using the regional interactions, and
12 show significant co-phylogenetic matching using the local interactions (see
*Supp. Mat. 1* for network-level significance values; @tbl:contingency;
@fig:maps). These results support the idea that macro-evolutionary processes,
such as co-diversification, can have consequences at the macro-ecological level
but may not in fact be detectable at finer spatial scales.

```{r congruence, results='markup', echo=F}

# Local structure or no?
local <- fig1dat$loc < 0.05
isloc <- local
isloc[local] <- "Local significant"
isloc[!local] <- "Local non-significant"

# Regional structure or no?
regal <- fig1dat$reg < 0.05
isreg <- regal
isreg[regal] <- "Regional significant"
isreg[!regal] <- "Regional non-significant"

kable(table(isloc, isreg))
```
Table: Contingency table showing the number of networks with (non-)significant local and regional co-phylogenetic signal. {#tbl:contingency}

## Local and regional scale networks have the same relative co-phylogenetic matching

```{r localregional, fig.show='hide', dev='pdf', echo=F, results='hide', warning=F, message=F, error=F}

z = function(x){return((x-mean(x))/sd(x))}
plot(z(loc_ss)~z(reg_ss), fig1dat, pch=19,
  xlab="Relative co-phylogenetic signal (regional)", ylab="Relative co-phylogenetic signal (local)", asp=1)
abline(h=0, lty=2, col="grey")
abline(v=0, lty=2, col="grey")
abline(a=0, b=1, col="darkgrey")
```

![The regional and local networks show the same relative amount of co-phylogenetic matching. The values presented are the z-scores of the PACo statistic for the entire network, with the 1:1 relationship indicated by the solid line.](figure/localregional-1.pdf){#fig:relative width=80%}

When we compared the relative degree of co-phylogenetic matching in the local
and regional communities ({@fig:relative}), we see that the relationship between
the two is approximately linear (95% confidence interval for the correlation
coefficient $0.914$--$0.971$). This fits with the hypothesis of local networks
being assembled by a random sampling from regional networks: if interactions
between species at matching positions in both trees are maintained by the same
set of drivers, then this should be reflected in the local networks by a higher
degree of co-phylogenetic matching.

## Co-phylogenetic matching is predicted by the contribution of interactions

```{r contributions, fig.show='hide', dev='pdf', echo=F, results='hide', warning=F, message=F, error=F, fig.width=10, fig.height=6}
library(RColorBrewer)
library(plyr)
palette(brewer.pal(8, "Set1"))

load("data/paco_fig2.Rdata")
load("data/paco_fig1.Rdata")
load("data/D.Rdata")

par(mfcol=c(1,2), xaxs="i", yaxs="i")

LOC <- fig1dat$loc_ss / max(D$gof$ss)
REG <- fig1dat$reg_ss / max(D$gof$ss)
imean = ddply(fig2dat[,c("site", "type", "rvalue")], .(site, type), summarize, m = mean(rvalue))
mloc = subset(imean, type=="loc")[,c("site", "m")]
colnames(mloc) = c("site", "mloc")
mreg = subset(imean, type!="loc")[,c("site", "m")]
colnames(mreg) = c("site", "mreg")
mi = merge(mloc, mreg)
colnames(mi) = c(".id", "mloc", "mreg")
allm = merge(mi, fig1dat)

plot(REG~mreg, allm, pch=NA, col=2, ylab = "Relative amount of co-phylogenetic signal", xlab="Average interaction-level contribution to co-phylogenetic signal", ylim=c(0, 0.3), xlim=rev(range(mreg)))
segments(x0=allm$mreg, x1=allm$mloc, y0=REG, y1=LOC, col="lightgrey")
lines(smooth.spline(x=allm$mreg, y=REG, df=3), lty=2, col=2)
lines(smooth.spline(x=allm$mloc, y=LOC, df=3), lty=2, col=3)
points(REG~mreg, allm, pch=19, col=2)
points(LOC~mloc, allm, pch=19, col=3)
legend("topright", fill=c(2,3), legend=c("Regional", "Local"), bty="n")
legend("topleft", pch=NA, legend=c("a"), bty="n", text.font=2, cex=1.2)
co = col2rgb(palette())/255
plot(density(unique(fig2dat[,c("int","rvalue")])$rvalue, from=1e3, to=max(fig2dat$rvalue)),
   xlab = "Interaction-level contribution to co-phylogenetic signal",
   ylab = "Probability density",
   col = NA,
   xlim = rev(c(0, 3e4)),
   ylim = c(0, 0.0002),
   main=""
)
hist(fig2dat$rvalue, col=rgb(co[1,1], co[2,1], co[3,1], alpha=0.3), add=T, freq=F, border=NA)
hist(subset(fig2dat, type=="reg")$score, col=rgb(co[1,2], co[2,2], co[3,2], alpha=0.3), add=T, freq=F, border=NA)
hist(subset(fig2dat, type=="loc")$score, col=rgb(co[1,3], co[2,3], co[3,3], alpha=0.3), add=T, freq=F, border=NA)
lines(density(unique(fig2dat[,c("int","rvalue")])$rvalue, from=0, to=max(fig2dat$rvalue)), lwd=2, col=1)
lines(density(subset(fig2dat, type=="reg")$score, from=0, to=max(fig2dat$rvalue)),col=2, lwd=2)
lines(density(subset(fig2dat, type=="loc")$score, from=0, to=max(fig2dat$rvalue)),col=3, lwd=2)
box()
legend("topright", fill=c(1,2,3), legend=c("Continental", "Regional", "Local"), bty="n")
legend("topleft", pch=NA, legend=c("b"), bty="n", text.font=2, cex=1.2)
```

![Distribution of co-phylogenetic matching at the network and interaction levels. **a**, Networks that have lower co-phylogenetic matching at the local or regional level are composed of interactions that on average contribute little to co-phylogenetic matching at the continental scale. Co-phylogenetic matching is presented relatively to the continental level co-phylogenetic matching. Dashed lines are a cubic smoothing spline, and the two levels of the same networks are linked by solid grey lines. **b**, Overall, interactions observed at the local, regional, and continental scale have roughly equivalent contributions to co-phylogenetic matching. Probability density was smoothed using a Gaussian kernel density estimator. Raw probability densities are shown as semi-transparent bars. Interaction-level contributions are unitless, and lower values represent stronger signal.](figure/contributions-1.pdf){#fig:contributions width=100%}

On the other hand, system-level differences say little about the behavior of
individual interactions. Despite the fact most coevolutionary mechanisms act at
the interaction level [@thom99rmc], most *measures* of it are expressed at the
community level. We observe here that networks with interactions that are
important for co-phylogenetic matching at the continental scale are also
important for co-phylogenetic matching at the local and regional scales as well
($\rho = 0.95$; {@fig:contributions}A). Intriguingly, we also find that the
distribution of individual interactions' contributions to co-phylogenetic
matching is strongly conserved, regardless of the scale at which the
interactions are quantified ({@fig:contributions}B). Because interactions differ
between each other in terms of their total contribution to co-phylogenetic
matching, this implies that their distribution across networks (*i.e.* whether
the local network contains a sample of strongly contributing, or weakly
contributing, interactions) is what actually drives differences in overall
co-phylogenetic matching. As such, network-level co-phylogenetic matching
emerges directly from the properties of interactions and is not a property of
the network itself.

## Interactions contributing to co-phylogenetic matching are marginally more spatially consistent

```{r consistency, fig.show='hide', dev='pdf', echo=F, results='hide', warning=F, message=F, error=F, fig.width=9}
load("data/paco_fig3.Rdata")

fig3dat$vary = cut(fig3dat$proba, breaks=seq(0.0, 1.0, 0.2))

boxplot(stat~vary, fig3dat, ylim=rev(c(0,3e4)), lty=1, pch=19, xlab="Interaction spatial consistency", ylab="Interaction-level contribution to co-phylogenetic signal")
points(jitter(match(fig3dat$vary, levels(fig3dat$vary))), fig3dat$stat, pch=21, cex=fig3dat$weight/7, bg=rgb(0, 0, 0, alpha=0.2), col=NA)
```

![Spatial consistency of an interaction and its contribution to co-phylogenetic matching. Note that because *PACo* gives low scores to interactions with a strong contribution to co-phylogenetic matching, the y-axis is reversed. Spatial consistency is defined as the probability of observing an interaction between two species given that they were observed to co-occur (the size of the point is proportional to the number of times an interaction was observed). Although statistically significant, there was no biologically meaningful relationship between spatial consistency and an interaction's importance for co-phylogenetic matching in the continental network ($R^2 \approx 0.01$, $\rho = -0.1$, $p \leq 10^{-5}$).](figure/consistency-1.pdf){#fig:consistency width=80%}

Beyond their contribution to co-phylogenetic matching, interactions also
ultimately differ in how frequently they vary when the species involved co-occur
[@olit15staa; @troj15gvm; @cars14bdp]. This can happen, for example, when one of
the partners is able to forage for optimal resources [@bett15prk]. Once more,
the literature on host-parasite interactions assumes that the reason why some
interactions are more frequent is because they reflect a significant past
history of coevolution [@guim07iia; @mora10bhi]; that is, the ecological
constraints emerge from evolutionary conservatism. Using a weighted Pearson's
correlation between the interaction frequency, interaction contribution to
co-phylogenetic matching, and the number of observations of each interaction as
the weight, we observe that this is marginally true ($\rho \approx -0.11$. $t
\approx -5.09$ with weights; $\rho \approx -0.10$, $t \approx -4.6$ without;
both significant at $\alpha = 0.05$; {@fig:consistency}). Recall that the
*negative* correlation here arises from the fact that high interaction-level
values in PACo means *low* contribution to co-phylogenetic signal. Nevertheless,
the significance of this result ought to be tempered by the fact that the $R^2$
of both regressions is close to $0.01$. Consequently, the association between
spatial consistency and contribution to co-phylogenetic signal, while
statistically significant, explains so little variance of either quantities that
it is likely of negligible biological importance. This implies that the spatial
consistency of an interaction does not necessarily reflect its evolutionary
past, but rather (possibly) extant ecological processes.

## The contribution of interactions to co-phylogenetic matching is consistent across scales

```{r scales, fig.show='hide', dev='pdf', echo=F, results='hide', warning=F, message=F, error=F, fig.width=9}
library(RColorBrewer)
library(plyr)

load("data/paco_fig1.Rdata")
load("data/paco_fig2.Rdata")

colnames(fig1dat)[1] = "site"
fig1dat$loc = fig1dat$loc < 0.05
fig1dat$reg = fig1dat$reg < 0.05
fig2dat = merge(fig1dat, fig2dat)
fig2dat$int <- as.vector(fig2dat$int)
fig2dat <- subset(fig2dat, type=="loc")
corpval = function(x, y) cor.test(x, y)$p.value
f2d <- ddply(fig2dat, .(site, loc, reg), summarize, cor = cor(score, rvalue), l = length(int), p = corpval(score, rvalue))
f2d$type = factor(paste(f2d$reg, f2d$loc))
levels(f2d$type) = c("None", "Local only", "Regional only", "Both")
par(xaxs="i", yaxs="i")

palette(brewer.pal(8, "Set2"))

plot(rank(f2d$p, ties.method="random"), f2d$cor, cex=log(f2d$l, 10), col="darkgrey", bg=as.numeric(f2d$type), xlim=c(0, 52), ylim=c(-1, 1), pch=NA, axes=T, ylab="Correlation between local and continental contribution to coevolutionary signal", xlab=NA)
rect(49.5, -1, 52, 1, col=rgb(0.88, 0.88, 0.82), border=NA)
rug(f2d$cor, side=2, col="grey")
rug(f2d$cor, side=4, col="grey")
abline(h=0, col="lightgrey", lwd=2, lty=2)
points(rank(f2d$p, ties.method="random"), f2d$cor, cex=log(f2d$l, 10), col="darkgrey", bg=as.numeric(f2d$type), xlim=c(0, 52), ylim=c(-1, 1), pch=21)

legend(x=-1, y=-0.5, legend=levels(f2d$type), fill=c(1,2,3,4), bty='n', title="Network coevolutionary signal", title.adj=7, xjust=0)
legend(x=-1.25, y=-0.1, legend=c(10, 30, 100), pch=19, bty='n', title="Number of interactions", title.adj=8, pt.cex=log(c(10, 30, 100),10))

box()
```

![The contribution to co-phylogenetic matching of the interaction between two species is maintained across scales. For every site (ranked on the x-axis), we show the Pearson's correlation between interaction-level values of co-phylogenetic matching in the continental network and the same in the local network. The size of each point is proportional to the size of the network, and correlations for all sites are significant at $\alpha = 0.05$ except for those falling in the grey shaded area.](figure/scales-1.pdf){#fig:scales width=100%}

Ultimately, co-phylogenetic matching varies across scale because of the
simultaneous variation of species' interactions *and* communities' phylogenetic
tree structure. In a system characterized by substantial turnover, we would
expect the contribution of each separate interaction to differ across scales as
well. Instead, we observe here that interactions that contribute strongly to
co-phylogenetic matching at the continental scale *also* show a significant
tendency to contribute strongly at the local ($p < 0.05$ for positive
correlations in 48 out of 51 networks) and regional (in 47 out of 51 networks),
and this observation is independent of network-wide co-phylogenetic matching
({@fig:scales}). Remarkably, this result implies that the remnants of
co-phylogenetic inertia are still locally detectable in *individual
interactions* even though shared evolutionary history regularly fails to leave
its imprint on most local networks.

# Conclusions

Overall, the results of our study demonstrate that there is a sizeable gap
between our current understanding of host-parasite coevolution as the basis of
multi-species interactions, its phylogenetic consequences, and their
applicability to ecological questions. Our results suggest that, while the
continental-scale system might show a strong signal of past coevolution through
significantly matching phylogenies (which was also reported, through different
analyses, by other studies of this system), the quasi-entirety of this signal is
lost when species and their interactions are filtered to assemble local
communities. That there is no further loss of signal from the regional to the
local scale strongly suggests that the loss of signal from the continental to
regional scale is due to species sampling in a manner that acts
independently of the evolutionary history of species pairs. Because regional and
local networks have the same species, the difference between them stems from the
loss of some interactions locally. It would therefore seem that local
species pools in this system are also driven by the interaction between abiotic
conditions and species tolerance, in addition to potential species interactions.
Taking a step back, this result suggests that, while a shared phylogenetic
history is a strong structuring force at the scale of the species pool, its
influence is overridden by other factors during species filtering and community
assembly. It further highlights the need for future investigation of whether the importance of
phylogenetic history in host-parasite
assemblages decays at smaller spatial scales.

Recently, @coel17nbp studied the factors that shape phylogenetic signal in
bipartite interactions between plants and mutualists. In contrast to our study,
they used a Mantel test to compare the two trees, a method whose lack of
robustness [@harm10psp; @guil13dmt] is what originally prompted the
development of PACo [@balb13pnp] or ParaFit [@lege02sth]. Setting aside the
relative merits of the various methods, they showed that phylogenetic signal is
expected to decrease when (1) dispersal probability increases (*i.e.* all
locations share the same species) and (2) the probability of interactions is
high (*i.e.* there is no loss of signal due to local interaction filtering). We
argue here that the probability of interaction should not be viewed as a system-wide
measure, but is better defined for each interaction separately ({@eq:linktunover}). When
this is the case, we report no relationship between the probability of
interaction and its contribution to co-phylogenetic signal ({@fig:consistency}).

The lack of a relationship between an interaction's probability of occurring (which represents
their ecological stability) and its contribution to the co-phylogenetic process
(which can be a proxy of their evolutionary stability) draws into question
the basic long-term stability of interactions. Specifically, the existence of
interactions with strong co-phylogenetic contribution, but low probability, is
puzzling. It is tempting to hypothesize that these interactions may have once been strong
enough to drive (co-)speciation, but are now decaying, due either to the fact
that the host-parasite pair has been disrupted or as a consequence of
ecological processes. DBS Can you spell this transition out for me better? I don't quite see the link. DBS @janz11err show that parasites can frequently acquire
hosts that are not closely related to their initial host. This phenomenon,
dubbed the "parasite paradox", can be responsible for a large number of
diversification events, as parasites will be introduced to a novel environment
when exposed to a novel host. The acquisition of novel hosts proceeds primarily
from ecological contact between distantly related species. This emphasizes the
need to account for the spatial structure of species interactions in these
studies -- because all species do not co-occur at all places, the potential for
ecological contact with novel hosts varies across space. Ideally, reconstruction
of the ancestral range of the species can shed some light on whether acquisition
of distantly related species happened, and where.

Local networks show little to no signal of co-phylogenetic matching, and the
strength of co-phylogenetic matching that can be ascribed to the interactions
between two species is a surprisingly poor predictor of how frequently they
interact. In contrast to the frequent assumption that phylogenetic structure is
a key driver of community structure [@cave09mce], these data reveal that this
impact is actually minimal at ecologically relevant spatial scales, even in
a system where phylogenetic structure would be expected to have a profoundly
strong role. And yet, despite all the above, individual interactions are somehow
able to maintain their co-phylogenetic matching even when the community they are
woven into does not. Thinking more broadly, these discrepancies provide a 
roadmap for bridging the gap between our appreciation of the role of shared
evolutionary history and its empirically measurable outcomes: network structure
is the most parsimonious *mechanism* by which coevolution proceeds, not the
imprint potential coevolution leaves on ecological communities.

Finally, we would like to acknowledge some limitations in the data.
Host-parasite evolution usually follows a number of canonical coevolutionary
scenarios [@page03ttp]: true cospeciation, where the two speciation events are
simultaneous; pseudo- or delayed-cospeciations, where one speciation event
happens after the other; host shift, where the parasite acquires a new host
(and can subsequently lose the previous host, undergo speciation, or maintain its
ability to use the previous host); and host-range expansion, in which the parasite
acquires a novel host.
DBS Are we sure the differences between host shift and host-range expansion clear enough in the above? DBS
When the likelihood of both host and parasite extinction
events are added to these models, the variety of outcomes is considerable.
@dev07wch shows that host-switch-like events can result in matching trees, and
so one could argue that the continental-level signal of tree matching can emerge
from numerous host-shift events [although the results of @kras12psm do not
support this scenario]. Because separating true cospeciation events from all
other situations is difficult without time-calibrated trees, @dev13cvh suggest
that the number of cospeciation events is inflated by cophenetic methods like
*PACo* or *ParaFit* [although @lope01mpf reported an increased number of
cospeciations when accounting for time of speciation events]. This reinforces
the need for high-quality phylogeny data calibrated for communities in which the
interactions are known. Not only will this help us make more robust inference about
the evolution of networks [@eklo16pcf], it will also help us leverage information about
species interactions embedded within networks to better understand coevolutionary
dynamics.

**Acknowledgements.** We thank Juan Antonio Balbuena for discussions about the
*PACo* method, and members of the Stouffer and Tylianakis groups for comments
on an early draft of this manuscript. We thank Scott Nuismer for feedback. We
are indebted to Matt Hutchinson and Fernando Cagua for contributions to the
code of the `paco` R package. Funding to TP and DBS was provided by a Marsden
Fund Fast-Start grant (UOC-1101) and to DBS by a Rutherford Discovery
Fellowship, both administered by the Royal Society of New Zealand.

# References